---
title: "project_osna1"
output: html_document
date: "2024-10-30"
---
```{r}
# Install and load packages
install.packages("readxl")
library(readxl)

install.packages("naniar")
library(naniar)

install.packages("ggplot2")
library(ggplot2)
# install.packages(c("dplyr", "ggplot2", "tidyverse"))
# library(dplyr)
# library(ggplot2)
# library(tidyverse)
```

```{r}
# Load and view dataset
demograpgh_data <- read_excel("~/Documents/demograpgh_data.xlsx", sheet = "Sheet1")
View(demograpgh_data)
head(demograpgh_data)
```

```{r}
# Structure and type of columns in dataset
str(demograpgh_data)
```

``````{r}
# Summary statistics
summary(demograpgh_data)
``````


```````{r}
# Check for Missing Values
colSums(is.na(demograpgh_data))
`````

```````{r}
# Check for Missing Values
colSums(is.na(demograpgh_data))

gg_miss_var(demograpgh_data)
`````
```````{r}
# Check for duplicate rows
duplicates <- data[duplicated(demograpgh_data), ]
print(duplicates)
``````

```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(scales)

```

```````{r}
# Univariate analysis
# Plot for total population gender based in entire ditrict
# Prepare data for stacked bar chart
library(ggplot2)

# Create the gender data frame
gender_data <- data.frame(
  Gender = c("Male", "Female"),
  Count = c(sum(demograpgh_data$male_population), sum(demograpgh_data$female_population))
)

# Create the pie chart with counts
ggplot(gender_data, aes(x = "", y = Count, fill = Gender)) +
  geom_bar(stat = "identity") +
  coord_polar("y") +
  labs(title = "Gender Distribution of Population", fill = "Gender") +
  theme_void() +  # Clean up the plot
  geom_text(aes(label = Count), 
            position = position_stack(vjust = 0.5),  # Position labels in the middle of slices
            color = "white")  # Adjust text color for visibility
``````

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)  # Load the scales package

# Create a summary data frame for gender population
gender_data <- demograpgh_data %>%
  select(countyname, male_population, female_population) %>%
  pivot_longer(cols = c(male_population, female_population), names_to = "Gender", values_to = "Count")

# Grouped bar plot for gender
ggplot(gender_data, aes(x = countyname, y = Count, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Gender Population by County", x = "County", y = "Population Count") +
  scale_y_continuous(labels = comma) +  # Format y-axis to show normal numbers
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

````

```{r}
#density plot for population in counties
library(ggplot2)
library(scales)  # Load scales for formatting

ggplot(demograpgh_data, aes(x = population)) + 
  geom_density(fill = "blue", alpha = 0.5) + 
  labs(title = "Density Plot of Population", x = "Population", y = "Density") + 
  scale_x_continuous(labels = comma) +  # Format x-axis to show normal numbers
  scale_y_continuous(labels = comma) +  # Format y-axis to show normal numbers
  theme_minimal()

```
```{r}
# Density plot for male and female populations
ggplot(gender_data, aes(x = Count, fill = Gender)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density Plot of Male and Female Populations", 
       x = "Population Count", 
       y = "Density") +
  scale_x_continuous(labels = comma) +  # Format x-axis to show normal numbers
  scale_y_continuous(labels = comma) +  # Format y-axis to show normal numbers
  theme_minimal()
```

```{r}
ggplot(demograpgh_data, aes(x = reorder(countyname, medinc), y = medinc)) + 
  geom_bar(stat = "identity", fill = "orange") + 
  labs(title = "Median Income by County", x = "County", y = "Median Income") + 
  scale_y_continuous(labels = scales::comma) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
# Load necessary library
# Load necessary library
library(ggplot2)

# Define age columns
age_columns <- c('age_18_19', 'age_20_24', 'age_25_29', 'age_30_34', 
                 'age_35_44', 'age_45_54', 'age_55_64', 'age_65_74', 
                 'age_75_84', 'age_85over')

# Create histograms for each age group
for (column in age_columns) {
    p <- ggplot(demograpgh_data, aes_string(x = column)) +
        geom_histogram(bins = 30, fill = "blue", color = "black", alpha = 0.5) +
        labs(title = paste("Distribution of", column), x = column, y = "Frequency") +
        theme_minimal() +
        theme(text = element_text(size = 12))
    
    print(p)  # Display the plot
}


```
```{r}
# Assuming your dataset is already loaded as 'demograpgh_data'

# Summarize gender counts for Male and Female only
gender_data <- data.frame(
    Gender = c("Male", "Female"),
    Count = c(sum(demograpgh_data$voters_gender_m, na.rm = TRUE),
               sum(demograpgh_data$voters_gender_f, na.rm = TRUE))
)

# Create pie chart for gender distribution
ggplot(gender_data, aes(x = "", y = Count, fill = Gender)) +
    geom_bar(stat = "identity") +
    coord_polar("y") +
    labs(title = "Gender Distribution of Voters", fill = "Gender") +
    theme_void() +
    geom_text(aes(label = paste0(round(Count / sum(Count) * 100, 1), "%")),
              position = position_stack(vjust = 0.5)) +
    theme(legend.position = "bottom")  # Optional: adjust legend position


```

```{r}
# Load necessary library
library(corrplot)

# Select numeric columns for correlation
numeric_data <- demograpgh_data[, c("population", "male_population", "female_population", "medinc", "pov_rate", "total_reg")]

# Calculate correlation matrix
cor_matrix <- cor(numeric_data)

# Plot correlation matrix
corrplot(cor_matrix, method = "circle")

```

```{r}
library(ggplot2)

# Ensure pov_rate is in decimal form
demograpgh_data$pov_rate <- demograpgh_data$pov_rate / 100

ggplot(demograpgh_data, aes(x = medinc, y = pov_rate)) +
    geom_point(color = "blue", alpha = 0.6) +
    labs(title = "Median Income vs. Poverty Rate",
         x = "Median Income",
         y = "Poverty Rate (%)") +
    scale_x_continuous(labels = scales::comma) +  # Format x-axis
    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
    theme_minimal()

```

```{r}
library(ggplot2)

ggplot(demograpgh_data, aes(x = population, y = medinc)) +
  geom_point(color = "blue", alpha = 0.6) +
  labs(title = "Scatter Plot of Population vs. Median Income", 
       x = "Population", 
       y = "Median Income") +
  scale_x_continuous(labels = scales::comma) +  # Format x-axis to remove scientific notation
  scale_y_continuous(labels = scales::comma) +   # Format y-axis to remove scientific notation
  theme_minimal()


```


